Easy as Pi, Volume II: On the Goodput and Energy of Network Coding Schemes in the Raspberry Pi

### 2016-08-14 introduction.tex
Some ideas

% RAW INTRODUCTION: NEED OF DISTRIBUTED NETWORKS AND THE RASPBERRY PI
Due to the advent of the \ac{IoT}, approximately 50 billion devices
ranging from sensors to phones are expected to be connected through
data networks in relatively short period of time \cite{cisco2011forecast}.
This massive deployment requires the design and testing of new
distributed systems that permit to manage the amount of traffic
from the proposed services provided by these devices. Therefore,
development platforms that help to quickly deploy, analyze and
evaluate this type of settings are highly desirable for research.
The emergence of a powerful and inexpensive portable computer such
as the Raspberry Pi, running standard operating systems such as
Linux or Windows, allows to achieve this goal since it enables the
deployment of distributed systems in a rapid fashion at a low cost.
Also, the use of an operating system that has a large user community
permits to administrative these networks in a flexible, stable and
supported fashion.

@@

In this work, we provide detailed measurements of the goodput (processing
speed) and energy consumption of \ac{Raspi} models 1 and 2, when
performing network coding operations with different codecs based on \ac{RLNC}
such as: full dense \ac{RLNC}, multicore-enabled \ac{RLNC}, sparse \ac{RLNC}
and tunable sparse \ac{RLNC}. For these coding schemes, the encoder and implementations from Kodo now detect and exploit the \ac{SIMD}
capabilities of the Raspberry Pi. We also assess the \ac{Raspi}
performance excluding the effect of packet losses or delays in order to have a
description of the codes in terms of their parameters. To achieve this,
we perform a measurement campaign with the indicated coding schemes and
their parameters in our local Raspberry Pi testbed. Our measurements permit
us to characterize the goodput and energy consumption
of these devices showing that speeds of 800 Mbps and processing energy per bit
values of 0.1 nJ are possible. Our work is organized as follows.
Section~\ref{sec:schemes} defines the
coding schemes employed in our study. Later, in Section~\ref{sec:metrics}
we describe the considered metrics and methodology for performance comparison
of the codes deployed in the \ac{Raspi}. In Section~\ref{sec:measurements}, we
show the measurements in the \ac{Raspi} models of the mentioned metrics
providing full discussions about the observed operational regimes and effects.
Final conclusions and future work are reviewed in
Section~\ref{sec:conclusions}.

>>>>

% RAW INTRODUCTION: NEED OF DISTRIBUTED NETWORKS AND THE RASPBERRY PI
Due to the advent of the \ac{IoT}, approximately 50 billion devices
ranging from sensors to phones are expected to be connected through
data networks in relatively short period of time \cite{cisco2011forecast}.
This massive deployment requires the design and testing of new
distributed systems that permit to manage the amount of traffic
from the proposed services provided by these devices. Therefore,
development platforms that help to quickly deploy, analyze and
evaluate this type of settings are highly desirable for research.
The emergence of a powerful and inexpensive portable computer such
as the Raspberry Pi makes this possible by enabling system designers
to quickly deploy many customizable devices at the same time
for a low cost.

By running standard operating systems such as
Linux or Windows, it allows to achieve this goal since it enables the
deployment of distributed systems in a rapid fashion at a low cost.
Also, the use of an operating system that has a large user community
permits to administrative these networks in a flexible, stable and
supported fashion.

@@

In this work, we provide detailed measurements of the goodput (processing
speed) and energy consumption of \ac{Raspi} models 1 and 2, when
performing network coding operations with different codecs based on \ac{RLNC}
such as: full dense \ac{RLNC}, multicore-enabled \ac{RLNC}, sparse \ac{RLNC}
and tunable sparse \ac{RLNC}. For these coding schemes, the encoder and implementations from Kodo now detect and exploit the \ac{SIMD}
capabilities of the Raspberry Pi. We also assess the \ac{Raspi}
performance excluding the effect of packet losses or delays in order to have a
description of the codes in terms of their parameters. To achieve this,
we perform a measurement campaign with the indicated coding schemes and
their parameters in the Raspberry Pi platform. Our measurements permit
us to characterize the goodput and energy consumption
of these devices showing that speeds of 800 Mbps and processing energy per bit
values of 0.1 nJ are possible. Our work is organized as follows.
Section~\ref{sec:schemes} defines the
coding schemes employed in our study. Later, in Section~\ref{sec:metrics}
we describe the considered metrics and methodology for performance comparison
of the codes deployed in the \ac{Raspi}. In Section~\ref{sec:measurements}, we
show the measurements in the \ac{Raspi} models of the mentioned metrics
providing full discussions about the observed operational regimes and effects.
Final conclusions and future work are reviewed in
Section~\ref{sec:conclusions}.
